---
title: "National Parks Slides"
author: "John Oatney, Gwen Wooten, Tommy Fowler"
date: "`r Sys.Date()`"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

# Here are the packages we used:

library(tidyverse)

library(openintro) 

library(ggpubr)  
  
library(leaflet)

devtools::install_github("wmurphyrd/fiftystater")
  
  library(fiftystater)

  data("fifty_states")
  
  
# Here are the data sets:

library(readr)
species <- read_csv("Biodiversity/species.csv")

library(readr)
parks <- read_csv("Biodiversity/parks.csv")

library(readr)
states <- read_csv("Biodiversity/states.csv")
```

```{r Key Data Set Variations, include = FALSE}
speciesendanger <- species %>% 
  filter(!is.na(`Conservation Status`), Occurrence != "Not Confirmed") %>% 
  filter(!Abundance %in% c("Native", "Not Native"))

speciesgraph <- speciesendanger %>% 
 filter(!is.na(Abundance), Abundance != 'Unknown') %>% 
  group_by(Category, Abundance) %>% 
  summarize(total = n()) %>% 
  filter(total > 10) 

speciesabundance <- species %>% 
  filter(!is.na(Abundance)) %>% 
  filter(!Abundance %in% c("Native", "Not Native")) %>% 
  group_by(Abundance, `Conservation Status`) %>% 
  summarize(count = n()) %>% 
  filter(`Conservation Status` == "Species of Concern")

speciesnew <- species %>% 
  filter(Occurrence != "Not Confirmed", 
         !Abundance %in% c("Native", "Not Native"), 
         !`Conservation Status` %in% c("Under Review", "In Recovery", "Extinct")) %>% 
  mutate(`Conservation Status` = case_when(`Conservation Status` %in% c("Species of Concern", "Endangered", "Threatened", "Proposed Threatened", "Proposed Endangered") ~ "Species of Concern")) %>% 
  mutate(`Conservation Status` = replace_na(`Conservation Status`, "Not of Concern"))
```


## R Markdown

This is an R Markdown presentation. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.

## Slide with Bullets

- Bullet 1
- Bullet 2
- Bullet 3

## Slide with R Output

```{r cars, echo = TRUE}
summary(cars)
```

## Slide with Plot

```{r echo = FALSE, message = FALSE, warning = FALSE}

graph2 <- species %>% 
 filter(!is.na(Abundance), Abundance != 'Unknown') %>% 
  filter(Occurrence != "Not Confirmed") %>% 
   group_by(Category, Abundance, `Conservation Status`) %>% 
  summarize(total = n()) %>% 
  #mutate(str_replace_all(`Conservation Status`, "(Endagered | Threatened)", "Threatened or Endagered")) %>%
  mutate(`Conservation Status` = replace_na(`Conservation Status`, "Not Threatened or Endagered")) %>% 
  filter(total > 10) 


#GRAPHS START

 vasc <- ggplot(graph2) +
geom_col(aes(y = fct_reorder(Category, total), x=total, fill=`Conservation Status`), stat='identity') +
     labs(y = " ", x = "Total Count") +
 # Making the background
  theme(panel.background = element_rect(fill = "white", 
                                        colour = "white", 
                                        linewidth = 2, 
                                        linetype = "solid"),
        panel.grid.major = element_line(linewidth = 0.5, 
                                        linetype = 'solid',
                                        colour = "white"), 
        panel.grid.minor = element_line(linewidth = 0.25, 
                                        linetype = 'solid',
                                        colour = "white")) +
  
  # Changing the legend
  guides(fill = guide_legend(title.position = "top", 
                             title = "Level of Concern")) + 
  
  # Giving it a title
  labs(title = " ") +
  
  # Affecting text attributes
  theme(axis.text.y = element_text(color = "grey27"),
        axis.text.x = element_text(size = 10, color = "grey27"), 
        axis.ticks.y = element_blank(), 
        plot.title = element_text(face = "bold", 
                                  size = 17, 
                                  hjust = 0.5, 
                                  color = "grey27"), 
        plot.subtitle = element_text(face = "italic", 
                                     size = 10, 
                                     hjust = 0.5, 
                                     color = "grey27"), 
        plot.caption = element_text(face = "italic", 
                                    color = "grey27"),
        
        # Affecting legend attributes
        legend.position = "right", 
        legend.justification = "left",  
        legend.background = element_rect(fill ="white"), 
        legend.text = element_text(face="italic", 
                                   color = "grey27"), 
        legend.title = element_text(color = "grey27", 
                                    face = "bold")) +
  
  # Affecting axis text attributes
  theme(axis.title.y=element_text(colour="grey27", 
                                  face = "bold")) +
  theme(axis.title.x=element_text(colour="grey27", 
                                   face = "bold")) 
  

 
nonvasc <- ggplot(speciesgraph) +
geom_col(aes(y = fct_reorder(Category, total), fill=factor(Abundance, levels = c("Rare", "Uncommon", "Occasional", "Common", "Abundant")), x=total), stat='identity') +
  labs(y = " ", x = "Total Count") +
   # Making the background
  theme(panel.background = element_rect(fill = "white", 
                                        colour = "white", 
                                        linewidth = 2, 
                                        linetype = "solid"),
        panel.grid.major = element_line(linewidth = 0.5, 
                                        linetype = 'solid',
                                        colour = "white"), 
        panel.grid.minor = element_line(linewidth = 0.25, 
                                        linetype = 'solid',
                                        colour = "white")) +
  
  # Changing the legend
  guides(fill = guide_legend(title.position = "top", 
                             title = "Level of Concern")) + 
  
  # Giving it a title
  labs(title = " ") +
  
  # Affecting text attributes
  theme(axis.text.y = element_text(color = "grey27"),
        axis.text.x = element_text(size = 10, color = "grey27"), 
        axis.ticks.y = element_blank(), 
        plot.title = element_text(face = "bold", 
                                  size = 17, 
                                  hjust = 0.5, 
                                  color = "grey27"), 
        plot.subtitle = element_text(face = "italic", 
                                     size = 10, 
                                     hjust = 0.5, 
                                     color = "grey27"), 
        plot.caption = element_text(face = "italic", 
                                    color = "grey27"),
        
        # Affecting legend attributes
        legend.position = "right", 
        legend.justification = "left",  
        legend.background = element_rect(fill ="white"), 
        legend.text = element_text(face="italic", 
                                   color = "grey27"), 
        legend.title = element_text(color = "grey27", 
                                    face = "bold")) +
  
  # Affecting axis text attributes
  theme(axis.title.y=element_text(colour="grey27", 
                                  face = "bold")) +
  theme(axis.title.x=element_text(colour="grey27", 
                                   face = "bold")) 
  
  
  ggarrange(vasc, nonvasc, nrow = 2, labels = c("Including Vascular Plants", "Excluding Vascular Plants"))
```

