---
title: "National Parks Slides"
author: "John Oatney, Gwen Wooten, Tommy Fowler"
date: "`r Sys.Date()`"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

# Here are the packages we used:

library(tidyverse)

library(openintro) 

library(ggpubr)  
  
library(leaflet)

devtools::install_github("wmurphyrd/fiftystater")
  
  library(fiftystater)

  data("fifty_states")
  
  
# Here are the data sets:

library(readr)
species <- read_csv("Biodiversity/species.csv")

library(readr)
parks <- read_csv("Biodiversity/parks.csv")

library(readr)
states <- read_csv("Biodiversity/states.csv")
```

```{r Key Data Set Variations, include = FALSE}
speciesendanger <- species %>% 
  filter(!is.na(`Conservation Status`), Occurrence != "Not Confirmed") %>% 
  filter(!Abundance %in% c("Native", "Not Native"))

speciesgraph <- speciesendanger %>% 
 filter(!is.na(Abundance), Abundance != 'Unknown') %>% 
  group_by(Category, Abundance, `Conservation Status`) %>% 
  summarize(total = n()) %>% 
  filter(total > 10) 

speciesabundance <- species %>% 
  filter(!is.na(Abundance)) %>% 
  filter(!Abundance %in% c("Native", "Not Native")) %>% 
  group_by(Abundance, `Conservation Status`) %>% 
  summarize(count = n()) %>% 
  filter(`Conservation Status` == "Species of Concern")

speciesnew <- species %>% 
  filter(Occurrence != "Not Confirmed", 
         !Abundance %in% c("Native", "Not Native"), 
         !`Conservation Status` %in% c("Under Review", "In Recovery", "Extinct")) %>% 
  mutate(`Conservation Status` = case_when(`Conservation Status` %in% c("Species of Concern", "Endangered", "Threatened", "Proposed Threatened", "Proposed Endangered") ~ "Species of Concern")) %>% 
  mutate(`Conservation Status` = replace_na(`Conservation Status`, "Not of Concern"))
```


## R Markdown

This is an R Markdown presentation. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.

## Slide with Bullets

- Bullet 1
- Bullet 2
- Bullet 3

## Slide with R Output

```{r cars, echo = TRUE}
summary(cars)
```

## Slide with Plot

```{r echo = FALSE, message = FALSE, warning = FALSE}

graph2 <- speciesnew %>% 
 filter(!is.na(Abundance), Abundance != 'Unknown') %>% 
  filter(Occurrence != "Not Confirmed") %>% 
   group_by(Category, Abundance, `Conservation Status`) %>% 
  summarize(total = n()) %>% 
  mutate(`Conservation Status` = replace_na(`Conservation Status`, "Not Threatened or Endagered")) %>% 
  filter(total > 10) 

graph3 <- graph2 %>% 
  filter(Category != "Vascular Plant")


#GRAPHS START

ggplot(graph2) +
geom_col(aes(y = reorder(Category, total), x=total, fill=`Conservation Status`), stat='identity') +
     labs(y = " ", x = "Total Count") +
 # Making the background
  theme(panel.background = element_rect(fill = "white", 
                                        colour = "white", 
                                        linewidth = 2, 
                                        linetype = "solid"),
        panel.grid.major = element_line(linewidth = 0.5, 
                                        linetype = 'solid',
                                        colour = "white"), 
        panel.grid.minor = element_line(linewidth = 0.25, 
                                        linetype = 'solid',
                                        colour = "white")) +
  
  # Changing the legend
  guides(fill = guide_legend(title.position = "top", 
                             title = "Level of Concern")) + 
  
  # Giving it a title
  labs(title = " ") +
  
  # Affecting text attributes
  theme(axis.text.y = element_text(color = "grey27"),
        axis.text.x = element_text(size = 10, color = "grey27"), 
        axis.ticks.y = element_blank(), 
        plot.title = element_text(face = "bold", 
                                  size = 17, 
                                  hjust = 0.5, 
                                  color = "grey27"), 
        plot.subtitle = element_text(face = "italic", 
                                     size = 10, 
                                     hjust = 0.5, 
                                     color = "grey27"), 
        plot.caption = element_text(face = "italic", 
                                    color = "grey27"),
        
        # Affecting legend attributes
        legend.position = "right", 
        legend.justification = "left",  
        legend.background = element_rect(fill ="white"), 
        legend.text = element_text(face="italic", 
                                   color = "grey27"), 
        legend.title = element_text(color = "grey27", 
                                    face = "bold")) +
  
  # Affecting axis text attributes
  theme(axis.title.y=element_text(colour="grey27", 
                                  face = "bold")) +
  theme(axis.title.x=element_text(colour="grey27", 
                                   face = "bold")) 
  


```

# Leaflet slide

```{r echo = FALSE}
# Because of the nature of leaflet, we had to manually insert each of the National Parks by hand

m <- leaflet() %>%
  addTiles() %>%  
  addMarkers(lng=-68.21, lat=44.35, popup="Acadia National Park") %>%
  addMarkers(lat = 38.68, lng = -109.57, popup = "Arches National Park") %>%
  addMarkers(lat = 43.75, lng = -102.50, popup = "Badlands National Park") %>%
  addMarkers(lat = 29.25, lng = -103.25, popup = "Big Bend National Park") %>%
  addMarkers(lat = 25.65, lng = -80.08, popup = "Biscayne National Park") %>%
  addMarkers(lat = 38.57, lng = -107.72, popup = "Black Canyon of the Gunnison National Park") %>%
  addMarkers(lat = 37.57, lng = -112.18, popup = "Bryce Canyon National Park") %>%
  addMarkers(lat = 38.20, lng = -109.93, popup = "Canyonlands National Park") %>%
  addMarkers(lat = 38.20, lng = -111.17, popup = "Capital Reef National Park") %>%
  addMarkers(lat = 32.17, lng = -104.44, popup = "Carlsbad Caverns National Park") %>%
  addMarkers(lat = 34.01, lng = -119.42, popup = "Channel Islands National Park") %>%
  addMarkers(lat = 33.78, lng = -80.78, popup = "Congaree National Park") %>%
  addMarkers(lat = 42.94, lng = -122.10, popup = "Crater Lake National Park") %>%
  addMarkers(lat = 41.24, lng = -81.55, popup = "Cuyahoga Valley National Park") %>%
  addMarkers(lat = 63.33, lng = -150.50, popup = "Denali National Park and Preserve") %>%
  addMarkers(lat = 36.24, lng = -116.82, popup = "Death Valley National Park") %>%
  addMarkers(lat = 24.63, lng = -82.87, popup = "Dry Tortugas National Park") %>%
  addMarkers(lat = 25.32, lng = -80.93, popup = "Everglades National Park") %>%
  addMarkers(lat = 67.78, lng = -153.30, popup = "Gates of the Arctic National Park and Preserve") %>%
  addMarkers(lat = 48.80, lng = -114.00, popup = "Glacier National Park") %>%
  addMarkers(lat = 58.50, lng = -137.00, popup = "Glacier Bay National Park and Preserve") %>%
  addMarkers(lat = 38.98, lng = -114.30, popup = "Great Basin National Park") %>%
  addMarkers(lat = 36.06, lng = -112.14, popup = "Grand Canyon National Park") %>%
  addMarkers(lat = 37.73, lng = -105.51, popup = "Great Sand Dunes National Park and Preserve") %>%
  addMarkers(lat = 35.68, lng = -83.53, popup = "Great Smoky Mountians National Park") %>%
  addMarkers(lat = 43.73, lng = -110.80, popup = "Grand Teton National Park") %>%
  addMarkers(lat = 31.92, lng = -104.87, popup = "Guadalupe Mountians National Park") %>%
  addMarkers(lat = 20.72, lng = -156.17, popup = "Haleakala Natioanl Park") %>%
  addMarkers(lat = 19.38, lng = -155.20, popup = "Hawaii Volcanoes National Park") %>%
  addMarkers(lat = 34.51, lng = -93.05, popup = "Hot Springs National Park") %>%
  addMarkers(lat = 48.10, lng = -88.55, popup = "Isle Royale National Park") %>%
  addMarkers(lat = 33.79, lng = -115.90, popup = "Joshua Tree National Park") %>%
  addMarkers(lat = 58.50, lng = -155.00, popup = "Katmai National Park and Preserve") %>%
  addMarkers(lat = 59.92, lng = -149.65, popup = "Kenai Fjords National Park") %>%
  addMarkers(lat = 67.55, lng = -159.28, popup = "Kobuk Valley National Park") %>%
  addMarkers(lat = 60.97, lng = -153.42, popup = "Lake Clark National Park and Preserve") %>% 
  addMarkers(lat = 40.49, lng = -121.51, popup = "Lassen Volcanic National Park") %>%
  addMarkers(lat = 37.18, lng = -86.10, popup = "Mammoth Cave National Park") %>%
  addMarkers(lat = 37.18, lng = -108.49, popup = "Mesa Verde National Park") %>%
  addMarkers(lat = 46.85, lng = -121.75, popup = "Mount Rainier National Park") %>%
  addMarkers(lat = 48.70, lng = -121.20, popup = "North Cascades National Park") %>%
  addMarkers(lat = 47.97, lng = -123.50, popup = "Olympic National Park") %>%
  addMarkers(lat = 35.07, lng = -109.78, popup = "Petrified Forest National Park") %>%
  addMarkers(lat = 36.48, lng = -121.16, popup = "Pinnacles National Park") %>%
  addMarkers(lat = 41.30, lng = -124.00, popup = "Redwood National Park") %>%
  addMarkers(lat = 40.40, lng = -105.58, popup = "Rocky Mountain National Park") %>%
  addMarkers(lat = 32.25, lng = -110.50, popup = "Saguaro National Park") %>%
  addMarkers(lat = 36.43, lng = -118.68, popup = "Sequoia and Kings Canyon National Parks") %>%
  addMarkers(lat = 38.53, lng = -78.35, popup = "Shenandoah National Park") %>%
  addMarkers(lat = 46.97, lng = -103.45, popup = "Theodore Roosevelt National Park") %>%
  addMarkers(lat = 48.50, lng = -92.88, popup = "Voyageurs National Park") %>%
  addMarkers(lat = 43.57, lng = -103.48, popup = "Wind Cave National Park") %>%
  addMarkers(lat = 61.00, lng = -142.00, popup = "Wrangell - St Elias National Park and Preserve") %>%
  addMarkers(lat = 44.60, lng = -110.50, popup = "Yellowstone National Park") %>%
  addMarkers(lat = 37.83, lng = -119.50, popup = "Yosemite National Park") %>%
  addMarkers(lat = 37.30, lng = -113.05, popup = "Zion National Park") 
  
m  # Print the map
```  

