---
title: "Species of Concern in National Parks: Assessing Conservation Danger Across the United States"
author: "John Oatney, Gwen Wooten, Tommy Fowler"
date: "`r Sys.Date()`"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

# Here are the packages we used:

library(tidyverse)

library(openintro) 

library(ggpubr)  
  
library(leaflet)

devtools::install_github("wmurphyrd/fiftystater")
  
  library(fiftystater)

  data("fifty_states")
  
  
# Here are the data sets:

library(readr)
species <- read_csv("Biodiversity/species.csv")

library(readr)
parks <- read_csv("Biodiversity/parks.csv")

library(readr)
states <- read_csv("Biodiversity/states.csv")
```

```{r Key Data Set Variations, include = FALSE}
speciesendanger <- species %>% 
  filter(!is.na(`Conservation Status`), Occurrence != "Not Confirmed") %>% 
  filter(!Abundance %in% c("Native", "Not Native"))

speciesgraph <- speciesendanger %>% 
 filter(!is.na(Abundance), Abundance != 'Unknown') %>% 
  group_by(Category, Abundance, `Conservation Status`) %>% 
  summarize(total = n()) %>% 
  filter(total > 10) 

speciesabundance <- species %>% 
  filter(!is.na(Abundance)) %>% 
  filter(!Abundance %in% c("Native", "Not Native")) %>% 
  group_by(Abundance, `Conservation Status`) %>% 
  summarize(count = n()) %>% 
  filter(`Conservation Status` == "Species of Concern")

speciesnew <- species %>% 
  filter(Occurrence != "Not Confirmed", 
         !Abundance %in% c("Native", "Not Native"), 
         !`Conservation Status` %in% c("Under Review", "In Recovery", "Extinct")) %>% 
  mutate(`Conservation Status` = case_when(`Conservation Status` %in% c("Species of Concern", "Endangered", "Threatened", "Proposed Threatened", "Proposed Endangered") ~ "Species of Concern")) %>% 
  mutate(`Conservation Status` = replace_na(`Conservation Status`, "Not of Concern"))
```


## National Parks in the US

```{r VIZ 1, echo = FALSE, message = FALSE, warning = FALSE}

# Because of the nature of leaflet, we had to manually insert each of the National Parks by hand

m <- leaflet() %>%
  addTiles() %>%  
  addMarkers(lng=-68.21, lat=44.35, popup="Acadia National Park") %>%
  addMarkers(lat = 38.68, lng = -109.57, popup = "Arches National Park") %>%
  addMarkers(lat = 43.75, lng = -102.50, popup = "Badlands National Park") %>%
  addMarkers(lat = 29.25, lng = -103.25, popup = "Big Bend National Park") %>%
  addMarkers(lat = 25.65, lng = -80.08, popup = "Biscayne National Park") %>%
  addMarkers(lat = 38.57, lng = -107.72, popup = "Black Canyon of the Gunnison National Park") %>%
  addMarkers(lat = 37.57, lng = -112.18, popup = "Bryce Canyon National Park") %>%
  addMarkers(lat = 38.20, lng = -109.93, popup = "Canyonlands National Park") %>%
  addMarkers(lat = 38.20, lng = -111.17, popup = "Capital Reef National Park") %>%
  addMarkers(lat = 32.17, lng = -104.44, popup = "Carlsbad Caverns National Park") %>%
  addMarkers(lat = 34.01, lng = -119.42, popup = "Channel Islands National Park") %>%
  addMarkers(lat = 33.78, lng = -80.78, popup = "Congaree National Park") %>%
  addMarkers(lat = 42.94, lng = -122.10, popup = "Crater Lake National Park") %>%
  addMarkers(lat = 41.24, lng = -81.55, popup = "Cuyahoga Valley National Park") %>%
  addMarkers(lat = 63.33, lng = -150.50, popup = "Denali National Park and Preserve") %>%
  addMarkers(lat = 36.24, lng = -116.82, popup = "Death Valley National Park") %>%
  addMarkers(lat = 24.63, lng = -82.87, popup = "Dry Tortugas National Park") %>%
  addMarkers(lat = 25.32, lng = -80.93, popup = "Everglades National Park") %>%
  addMarkers(lat = 67.78, lng = -153.30, popup = "Gates of the Arctic National Park and Preserve") %>%
  addMarkers(lat = 48.80, lng = -114.00, popup = "Glacier National Park") %>%
  addMarkers(lat = 58.50, lng = -137.00, popup = "Glacier Bay National Park and Preserve") %>%
  addMarkers(lat = 38.98, lng = -114.30, popup = "Great Basin National Park") %>%
  addMarkers(lat = 36.06, lng = -112.14, popup = "Grand Canyon National Park") %>%
  addMarkers(lat = 37.73, lng = -105.51, popup = "Great Sand Dunes National Park and Preserve") %>%
  addMarkers(lat = 35.68, lng = -83.53, popup = "Great Smoky Mountians National Park") %>%
  addMarkers(lat = 43.73, lng = -110.80, popup = "Grand Teton National Park") %>%
  addMarkers(lat = 31.92, lng = -104.87, popup = "Guadalupe Mountians National Park") %>%
  addMarkers(lat = 20.72, lng = -156.17, popup = "Haleakala Natioanl Park") %>%
  addMarkers(lat = 19.38, lng = -155.20, popup = "Hawaii Volcanoes National Park") %>%
  addMarkers(lat = 34.51, lng = -93.05, popup = "Hot Springs National Park") %>%
  addMarkers(lat = 48.10, lng = -88.55, popup = "Isle Royale National Park") %>%
  addMarkers(lat = 33.79, lng = -115.90, popup = "Joshua Tree National Park") %>%
  addMarkers(lat = 58.50, lng = -155.00, popup = "Katmai National Park and Preserve") %>%
  addMarkers(lat = 59.92, lng = -149.65, popup = "Kenai Fjords National Park") %>%
  addMarkers(lat = 67.55, lng = -159.28, popup = "Kobuk Valley National Park") %>%
  addMarkers(lat = 60.97, lng = -153.42, popup = "Lake Clark National Park and Preserve") %>% 
  addMarkers(lat = 40.49, lng = -121.51, popup = "Lassen Volcanic National Park") %>%
  addMarkers(lat = 37.18, lng = -86.10, popup = "Mammoth Cave National Park") %>%
  addMarkers(lat = 37.18, lng = -108.49, popup = "Mesa Verde National Park") %>%
  addMarkers(lat = 46.85, lng = -121.75, popup = "Mount Rainier National Park") %>%
  addMarkers(lat = 48.70, lng = -121.20, popup = "North Cascades National Park") %>%
  addMarkers(lat = 47.97, lng = -123.50, popup = "Olympic National Park") %>%
  addMarkers(lat = 35.07, lng = -109.78, popup = "Petrified Forest National Park") %>%
  addMarkers(lat = 36.48, lng = -121.16, popup = "Pinnacles National Park") %>%
  addMarkers(lat = 41.30, lng = -124.00, popup = "Redwood National Park") %>%
  addMarkers(lat = 40.40, lng = -105.58, popup = "Rocky Mountain National Park") %>%
  addMarkers(lat = 32.25, lng = -110.50, popup = "Saguaro National Park") %>%
  addMarkers(lat = 36.43, lng = -118.68, popup = "Sequoia and Kings Canyon National Parks") %>%
  addMarkers(lat = 38.53, lng = -78.35, popup = "Shenandoah National Park") %>%
  addMarkers(lat = 46.97, lng = -103.45, popup = "Theodore Roosevelt National Park") %>%
  addMarkers(lat = 48.50, lng = -92.88, popup = "Voyageurs National Park") %>%
  addMarkers(lat = 43.57, lng = -103.48, popup = "Wind Cave National Park") %>%
  addMarkers(lat = 61.00, lng = -142.00, popup = "Wrangell - St Elias National Park and Preserve") %>%
  addMarkers(lat = 44.60, lng = -110.50, popup = "Yellowstone National Park") %>%
  addMarkers(lat = 37.83, lng = -119.50, popup = "Yosemite National Park") %>%
  addMarkers(lat = 37.30, lng = -113.05, popup = "Zion National Park") 
  
m  # Print the map

```


## Number of Plants and Animals 

```{r echo = FALSE, message = FALSE, warning = FALSE}
speciesnewabundance2 <- speciesnew %>% 
  filter(!is.na(`Abundance`), Abundance != 'Unknown') %>% 
  filter(Occurrence != "Not Confirmed") %>% 
  group_by(Category, Abundance, `Conservation Status`) %>% 
  summarize(total = n()) 
  
speciesnewabundance2 %>% group_by(Category) %>% 
  mutate(Grandtotal = sum(total)) %>% 
  ungroup() %>% # We want to redefine Category, so we have to ungroup it first
  mutate(Category = fct_reorder(Category, Grandtotal)) %>% 
  ggplot() + 
  geom_col(aes(y = Category, x = total)) +
  labs(y = " ", x = "Total Count") +

  # Hereafter are the theme details
  
  # Making the background
  theme(panel.background = element_rect(fill = "white", 
                                        colour = "white", 
                                        linewidth = 2, 
                                        linetype = "solid"),
        panel.grid.major = element_line(linewidth = 0.5, 
                                        linetype = 'solid',
                                        colour = "white"), 
        panel.grid.minor = element_line(linewidth = 0.25, 
                                        linetype = 'solid',
                                        colour = "white")) +
  
  # Affecting text attributes
  theme(axis.text.y = element_text(color = "grey27"),
        axis.text.x = element_text(size = 10, color = "grey27"), 
        axis.ticks.y = element_blank(), 
        plot.title = element_text(face = "bold", 
                                  size = 17, 
                                  hjust = 0.5, 
                                  color = "grey27"), 
        plot.subtitle = element_text(face = "italic", 
                                     size = 10, 
                                     hjust = 0.5, 
                                     color = "grey27"), 
        plot.caption = element_text(face = "italic", 
                                    color = "grey27")) + 
  
  # Affecting axis text attributes
  theme(axis.title.y=element_text(colour="grey27", 
                                  face = "bold")) +
  theme(axis.title.x=element_text(colour="grey27", 
                                   face = "bold"))

```  

## Research Question

Is there any state that needs more support in conserving its national parks' wildlife than others?

## Conservation Efforts

- The National Park Service
- Individual States
- Funding

## How Often Are Species Assigned Each Conservation Status?

```{r include=FALSE}

specieslargeviz <- speciesendanger %>% 
  filter(!is.na(Abundance)) %>% 
  group_by(Abundance, `Conservation Status`) %>% 
  summarize(count = n()) %>% 
  filter(!is.na(`Conservation Status`))

speciesconservationviz <- speciesendanger %>% 
  group_by(`Conservation Status`) %>% 
  summarize(count = n()) %>% 
  filter(!is.na(`Conservation Status`)) %>% 
  arrange(desc(count)) %>% 
  mutate(perc = (count / sum(count)) * 100) %>% 
  mutate(ypos = cumsum(perc)- 0.4*perc ) %>% 
  mutate(perc2 = round(perc, 1)) %>% 
  mutate(perc3 = paste0(perc2, "%"))

speciesabundanceviz <- speciesendanger %>% 
  filter(!is.na(Abundance)) %>% 
  group_by(Abundance) %>% 
  summarize(count = n())
```

```{r echo = FALSE, message = FALSE, warning = FALSE}
spc <- ggplot(speciesconservationviz, aes(x = "", y = perc, fill = reorder(`Conservation Status`, count), label = perc3)) +
  geom_bar(stat = "identity", width = 1, color = "white") + 
  coord_polar("y", start = 0) + 
  theme_void() + 
  scale_fill_manual(values = c("#8F0F1B", "#BE3ACD", "#304F9B", "#C7DD91", "#48A421", "#69EF7B", "#11CCDC", "black"),
                    breaks = c("Species of Concern", "Endangered", "Under Review", "Threatened", "In Recovery", "Proposed Endangered", "Proposed Threatened", "Extinct")) +
  geom_text(aes(y = ypos), color = "white", size = 6, check_overlap = TRUE) + 
    labs(fill = "Conservation Status", 
         title = " ")

spc
```

## How Much Concern Are Species Under?



```{r echo = FALSE, message = FALSE, warning = FALSE}
graph2 <- speciesnew %>% 
 filter(!is.na(Abundance), Abundance != 'Unknown') %>% 
  filter(Occurrence != "Not Confirmed") %>% 
   group_by(Category, Abundance, `Conservation Status`) %>% 
  summarize(total = n()) %>% 
  mutate(`Conservation Status` = replace_na(`Conservation Status`, "Not Threatened or Endagered"))


#GRAPHS START

ggplot(graph2) +
geom_col(aes(y = reorder(Category, total), x=total, fill=`Conservation Status`), stat='identity') +
     labs(y = " ", x = "Total Count") +
 # Making the background
  theme(panel.background = element_rect(fill = "white", 
                                        colour = "white", 
                                        linewidth = 2, 
                                        linetype = "solid"),
        panel.grid.major = element_line(linewidth = 0.5, 
                                        linetype = 'solid',
                                        colour = "white"), 
        panel.grid.minor = element_line(linewidth = 0.25, 
                                        linetype = 'solid',
                                        colour = "white")) +
  
  # Changing the legend
  guides(fill = guide_legend(title.position = "top", 
                             title = "Level of Concern")) + 
  
  # Giving it a title
  labs(title = " ") +
  
  # Affecting text attributes
  theme(axis.text.y = element_text(color = "grey27"),
        axis.text.x = element_text(size = 10, color = "grey27"), 
        axis.ticks.y = element_blank(), 
        plot.title = element_text(face = "bold", 
                                  size = 17, 
                                  hjust = 0.5, 
                                  color = "grey27"), 
        plot.subtitle = element_text(face = "italic", 
                                     size = 10, 
                                     hjust = 0.5, 
                                     color = "grey27"), 
        plot.caption = element_text(face = "italic", 
                                    color = "grey27"),
        
        # Affecting legend attributes
        legend.position = "right", 
        legend.justification = "left",  
        legend.background = element_rect(fill ="white"), 
        legend.text = element_text(face="italic", 
                                   color = "grey27"), 
        legend.title = element_text(color = "grey27", 
                                    face = "bold")) +
  
  # Affecting axis text attributes
  theme(axis.title.y=element_text(colour="grey27", 
                                  face = "bold")) +
  theme(axis.title.x=element_text(colour="grey27", 
                                   face = "bold")) 
  



```

## Birds as an Indicator of Concern

```{r include = FALSE}
speciespark <- parks %>% 
  full_join(species)

speciespark2 <- speciespark %>% 
filter(Occurrence != "Not Confirmed", !Abundance %in% c("Native", "Not Native"), !`Conservation Status` %in% c("Under Review", "In Recovery", "Extinct")) %>% 
  mutate(`Conservation Status` = case_when(`Conservation Status` %in% c("Species of Concern", "Endangered", "Threatened", "Proposed Threatened", "Proposed Endangered") ~ "Species of Concern")) %>% 
  mutate(`Conservation Status` = replace_na(`Conservation Status`, "Not of Concern")) 

concernpercent_perstate <- speciespark2 %>% 
  separate(State, c("State1", "State2", "State3"), sep = ", ") %>% 
  pivot_longer(cols = c(State1, State2, State3), names_to = "State Type", values_to = "State") %>% 
  filter(!is.na(State)) %>% 
  group_by(State, `Conservation Status`) %>% 
  summarize(total_ofconcern = n()) %>% 
  mutate(total_species_perstate = sum(total_ofconcern)) %>% 
  ungroup() %>% 
  mutate(percent = (total_ofconcern/total_species_perstate)*100) %>% 
  filter(`Conservation Status` != "Not of Concern")

parkstate <- speciesendanger %>% 
  group_by(`Park Name`, `Conservation Status`) %>% 
  summarize(count = n()) %>% 
  pivot_wider(names_from = `Conservation Status`, values_from = count, values_fill = 0) %>% 
  left_join(parks) %>% 
  select(`Park Code`, `Park Name`, State, Acres, Endangered, `In Recovery`, `Species of Concern`, `Under Review`, Threatened, `Proposed Threatened`, Extinct, `Proposed Endangered`) %>% 
  separate(State, c("State1", "State2", "State3"), sep = ", ") %>% 
  pivot_longer(cols = c(State1, State2, State3), names_to = "State Type", values_to = "State") %>% 
  filter(!is.na(State))

statevalues <- parkstate %>% 
  group_by(State) %>% 
  summarize(Concern = sum(`Species of Concern`), ConcernAvg = mean(`Species of Concern`), Count = n()) %>% 
  mutate(ConcernAvg = round(ConcernAvg, 2))

statevaluesmap <- statevalues %>% 
  full_join(states) %>% 
   mutate(state_name = str_to_lower(abbr2state(State)))

statevaluesmap_perc <- concernpercent_perstate %>% 
  full_join(states) %>% 
   mutate(state_name = str_to_lower(abbr2state(State)))

statecategories <- speciesnew %>% 
  filter(Category != "Vascular Plant") %>% 
  group_by(`Park Name`, Category) %>% 
  left_join(parks) %>% select(`Park Code`, `Park Name`, State, Acres, Category) %>% 
  separate(State, c("State1", "State2", "State3"), sep = ", ") %>% 
  pivot_longer(cols = c(State1, State2, State3), names_to = "State Type", values_to = "State") %>% 
  filter(!is.na(State)) %>% 
  group_by(State, Category) %>% 
  summarize(count = n()) %>% 
  
  # Now we do an inefficient but accurate method of pulling out the highest categories
  
  filter(count %in% c(1359, 300, 897, 2130, 811, 1225, 2149, 1564, 269, 243, 244, 230, 1754, 644, 159, 606, 242, 147, 473, 426, 649, 987, 712, 682, 1715, 359)) %>% 
  
  # Now we join it to make it work as a map
  
  full_join(states) %>% 
  mutate(state_name = str_to_lower(abbr2state(State)))
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
statecategories %>% 
  ggplot() +
  geom_map(
    map = fifty_states,
    aes(
      map_id = state_name,
      fill = Category
    ), color = "black"
  ) + 
  expand_limits(x = map_data("state")$long, y = map_data("state")$lat) + 
  labs(fill = "Most Common\nSpecies\nCategory") +
  theme_void()
```

## Birds as an Indicator of Concern

```{r echo = FALSE, message = FALSE, warning = FALSE}
statecategories %>% 
  filter(Category == "Bird") %>% 
  ggplot() + 
  geom_bar(
    aes(
      x = fct_reorder(State, count), 
      y = count
    ), stat = "identity"
  ) + 
  labs(x = "State", y = "# of Species", title = "How Many Bird Species Are There In States\nWhere Birds Are The Most Common Category?")
```

## How Is Each State Doing?

```{r VIZ 6, include= FALSE}
statevaluesmap <- statevalues %>% 
  full_join(states) %>% 
   mutate(state_name = str_to_lower(abbr2state(State)))

statevaluesmap_perc <- concernpercent_perstate %>% 
  full_join(states) %>% 
   mutate(state_name = str_to_lower(abbr2state(State)))
```


```{r echo = FALSE, message = FALSE, warning = FALSE}
#Average Number of Species of Concern per State

concernavg <- statevaluesmap %>%
  ggplot() +
  geom_map(
    map = fifty_states,
    aes(
      map_id = state_name,
      fill = ConcernAvg
    )
  ) +
  expand_limits(x = map_data("state")$long, y = map_data("state")$lat) + 
  labs(fill = "Avg. #\nSpecies of\nConcern") +
  theme_void()

#Total Number of Species of Concern per State

concerncount <- statevaluesmap %>%
  ggplot() +
  geom_map(
    map = fifty_states,
    aes(
      map_id = state_name,
      fill = Concern
    )
  ) +
  expand_limits(x = map_data("state")$long, y = map_data("state")$lat) + 
  labs(fill = "Total Species\nof Concern") +
  theme_void()

#Total Number of Parks per State

parkcount <- statevaluesmap %>%
  ggplot() +
  geom_map(
    map = fifty_states,
    aes(
      map_id = state_name,
      fill = Count
    )
  ) +
  expand_limits(x = map_data("state")$long, y = map_data("state")$lat) + 
  labs(fill = "# of\nParks") +
  theme_void()

#Percent of Species of Concern vs. Not of Concern per State

concernperc <- statevaluesmap_perc %>%
  ggplot() +
  geom_map(
    map = fifty_states,
    aes(
      map_id = state_name,
      fill = percent
    )
  ) +
  expand_limits(x = map_data("state")$long, y = map_data("state")$lat) + 
  labs(fill = "Percent\nSpecies of\nConcern") +
  theme_void()

ggarrange(
  parkcount, concerncount, concernavg, concernperc, ncol = 2, nrow = 2)

```


## Conclusion

```{r echo = FALSE, message = FALSE, warning = FALSE}

library(ggmap)
library(dplyr)

californiaconclusion <- read_csv("Biodiversity/californiaconclusion.csv")

caliconc <- californiaconclusion %>% 
  select(c('Park Code', 'Park', '# Endangered Species', 'Latitude', 'Longitude'))

US_map2 <- get_stamenmap(
  bbox = c(left = -125, bottom = 33, right = -115, top = 45),
  maptype = "terrain",
  zoom = 7
)

ggmap(US_map2) +
  geom_density_2d(data = caliconc, aes(x = Longitude, y = Latitude, ), size = 0.3) +
  stat_density_2d(data = caliconc, aes(x = Longitude, y = Latitude, fill = stat(level)), size = 0.05, alpha = .2, bins = 10, geom = "polygon", color = 'darkblue') +
  scale_alpha(guide = 'none') +
  scale_fill_gradient(low = "darkblue", high = "red", guide = 'none') +
  labs(title = "Density of Endangered Species\nby National Park in California") +
  theme(axis.title.y=element_blank()) +
  theme(axis.title.x= element_blank()) +
  theme(axis.ticks.y = element_blank()) +
  theme(axis.ticks.x = element_blank())


    
```

-